In this step, we already got many many models by filtering them using AUC.

What about their performance in outside data?

# Preparation
```{r library, include=FALSE}
source("Function.R")

tmp <- list()
```

Filtered pathway results:
```{r}
res_df <- readRDS("Output/Steps/R01/res_df_0.9.Rds")

pan_anno_df <- readRDS(file = "Output/Steps/R01/pan_anno.Rds")
res_df_sel <- res_df[res_df$Type %in% "UVM",]

eg2alias <- readRDS("Input/eg2alias_anndb.Rds")
# table(res_df$Type, res_df$method)
```

# Validate data preparation
## GSE84976
array:
```{r}
gset <- getGEO("GSE84976", destdir = "Input/Validation/")

exprSet1 <- exprs(gset[[1]])
exprSet1 <- as.data.frame(exprSet1)

GSE84976_clinical <- pData(gset[[1]])
table(GSE84976_clinical$`death due to metastasis:ch1`)
GSE84976_clinical$OS <- ifelse(
  GSE84976_clinical$`death due to metastasis:ch1` %in% "alive", 0, 1
)

# probe <- gset[[1]]@featureData@data
probe <- as.data.frame(fread("Input/Validation/GSE84976_family.txt"))
# probe <- probe[-1,]
probe <- probe[match(rownames(exprSet1), probe$ID),]
probe$Entrez_Gene_ID2 <- as.character(probe$Entrez_Gene_ID)

identical(probe$ID, rownames(exprSet1))
```

```{r}
GSE84976_exp <- exprSet1
GSE84976_exp$genes <- probe$Symbol

GSE84976_exp <- aggregate(.~genes,max,data=GSE84976_exp)
rownames(GSE84976_exp) <- GSE84976_exp$genes
GSE84976_exp$genes <- NULL

rownames(GSE84976_exp) <- gsub("-", "_", rownames(GSE84976_exp), fixed = T)
```

```{r}
GSE84976_clinical <- GSE84976_clinical[match(colnames(GSE84976_exp), rownames(GSE84976_clinical)),]
GSE84976_clinical$Source <- "GSE84976"
```


## GSE22138
array:
```{r}
gset2 <- getGEO("GSE22138", destdir = "Input/Validation/")

exprSet2 <- exprs(gset2[[1]])
exprSet2 <- as.data.frame(exprSet2)

GSE22138_clinical <- pData(gset2[[1]])
table(GSE84976_clinical$`death due to metastasis:ch1`)
GSE22138_clinical$OS <- ifelse(
  GSE22138_clinical$`metastasis:ch1` %in% "no", 0, 1
)

probe <- gset2[[1]]@featureData@data
# probe <- as.data.frame(fread("Input/Validation/GSE84976_family.txt"))
# probe <- probe[-1,]
probe <- probe[match(rownames(exprSet2), probe$ID),]
# probe$Entrez_Gene_ID2 <- as.character(probe$Entrez_Gene_ID)

identical(probe$ID, rownames(exprSet2))
```

```{r}
GSE22138_exp <- exprSet2
GSE22138_exp$genes <- probe$`Gene Symbol`

GSE22138_exp <- aggregate(.~genes,max,data=GSE22138_exp)
rownames(GSE22138_exp) <- GSE22138_exp$genes
GSE22138_exp$genes <- NULL

rownames(GSE22138_exp) <- gsub("-", "_", rownames(GSE22138_exp), fixed = T)
```

```{r}
GSE22138_clinical <- GSE22138_clinical[match(colnames(GSE22138_exp), rownames(GSE22138_clinical)),]
GSE22138_clinical$Source <- "GSE22138"
```

## Extra
```{r}
ggvenn::ggvenn(list(
  GSE84976 = rownames(GSE84976_exp),
  GSE22138 = rownames(GSE22138_exp)
))
```

```{r}
intersected_genes <- Reduce(intersect, list(rownames(GSE84976_exp),
                                            rownames(GSE22138_exp),
                                            rownames(meso_exp)))
```


# PCA test
```{r}
GSE84976_clinical$sample <- GSE84976_clinical$geo_accession
GSE22138_clinical$sample <- GSE22138_clinical$geo_accession

clinical <- rbind(
  GSE84976_clinical[,c(45,43,44)],
  GSE22138_clinical[,c(57,55,56)]
)
clinical$group <- ifelse(clinical$OS %in% 1, "Dead", "Alive")
```

```{r}
all_exp <- cbind(
  GSE84976_exp[match(intersected_genes,rownames(GSE84976_exp)),],
  GSE22138_exp[match(intersected_genes,rownames(GSE22138_exp)),]
)
```

```{r}
exp_pca <- PCA(t(all_exp), graph = F)
```

```{r}
fviz_pca_ind(exp_pca,
                             geom.ind = "point", # show points only (nbut not "text")
                             col.ind = factor(clinical$Source), # color by groups
                             addEllipses = F, palette = paletteer::paletteer_d("ggsci::default_igv")# Concentration ellipses
                             # legend.title = "Batch"
)

fviz_pca_ind(exp_pca,
                             geom.ind = "point", # show points only (nbut not "text")
                             col.ind = factor(clinical$group), # color by groups
                             addEllipses = F, palette = paletteer::paletteer_d("ggsci::default_igv")# Concentration ellipses
                             # legend.title = "Batch"
)
```

# Before Validation
## Tune_Grid making
Lasso:
```{r}
tune_grids_lasso <- separate(res_df_sel[res_df_sel$method %in% "lasso",c(4,7,8)], col = tune,
                             into = c("seed", "cv", "alpha"), sep = ":")
tune_grids_lasso$index <- 1:nrow(tune_grids_lasso)
tmp$tt <- sapply(1:4, function(x){tune_grids_lasso[[x]] <<- as.numeric(tune_grids_lasso[[x]])})

tune_grids_lasso <- as.data.frame(unique(tune_grids_lasso))
```

# Start Validate
## Function Define
```{r}
pathway_list <- readRDS("Output/Steps/R01/msigdb_selected_pathways2_company2.Rds")
tumor <- "UVM"
tumor_file <- paste0(tumor, ".Rds")
my_path <- "Output/Steps/R01/pan_RNA_exp/"
test_exp <- readRDS(paste0(my_path, tumor_file))
test_survival <- pan_anno_df[pan_anno_df$Project %in% tumor,]
```

## normal mode
```{r}
# GSE84976_val <- valLasso(tune_grids_lasso2, GSE84976_exp,
#                          GSE84976_clinical)
# meso_val <- valLasso(tune_grids_lasso2, meso_exp,
#                          meso_clinical)
# GSE22138_val <- valLasso(tune_grids_lasso2, GSE22138_exp,
#                          GSE22138_clinical)
```

## fast mode
```{r}
sfInit(parallel = TRUE, cpus = 6)
sfExport("test_exp", "test_survival", "pathway_list",
         "eg2alias")
sfExport('valProcess', 'symbol2Alias')
sfSource("Function.R")
```


```{r}
# tidyValOutput <- function(tune_re4){
#   tune_re2 <- do.call(c, tune_re4)
#   names(tune_re2) <- gsub("^.*]", "", names(tune_re2))
#   names(tune_re2) <- gsub(".", "", names(tune_re2), fixed = T)
#   tune_re2 <- data.frame(
#   AUC = tune_re2,
#   index = names(tune_re2)
#   )
#   return(tune_re2)
# }
```

The progress may get stuck for snowfall, you need to control the rows of data.frame. 

GSE84976:
```{r}
tune_list <- split(tune_grids_lasso, cut(seq_len(nrow(tune_grids_lasso)), 20))
tune_re <- lapply(tune_list, function(tune_grid){
  re <- valLasso_fast(tune_grid, GSE84976_exp,
                                GSE84976_clinical)
  print("finished")
  return(re)
})
```
```{r}
tune_re_1 <- do.call(c, tune_re)
tune_re_1 <- tune_re_1[tune_re_1 > .85]
```

GSE22138:
```{r}
tune_grids_lasso2 <- tune_grids_lasso[tune_grids_lasso$index %in% tune_re_1$index,]
tune_list <- split(tune_grids_lasso2, cut(seq_len(nrow(tune_grids_lasso2)), 2))
tune_re <- lapply(tune_list, function(tune_grid){
  re <- valLasso_fast(tune_grid, GSE84976_exp,
                                GSE84976_clinical)
  print("finished")
  return(re)
})
```
```{r}
tune_re_2 <- do.call(c, tune_re)
tune_re_2 <- tune_re_2[tune_re_2$AUC > .7,]
```


# select top model
```{r}
tune_all_res <- tune_grids_lasso[match(names(tune_re_2), tune_grids_lasso$index),]
tune_all_res$GSE84976 <- tune_re_1
tune_all_res$GSE22138 <- tune_re_2
```

```{r}
tune_all_res$AUC_all <- rowMeans(tune_all_res[,c("AUC",
                                                 "GSE84976",
                                                 "GSE22138")])
```

# Save
```{r}
saveRDS(tune_all_res,
     file = "Output/Steps/R02/validate_re.Rds")
```


external validation:
```{r}
save(GSE84976_exp, GSE84976_clinical,
     GSE22138_exp, GSE22138_clinical,
     file = "Output/Steps/R03/externals.Rda")
```

